Notes on Semi-Supervised `Novelty Detection'

### 2011-04-14 semisupervised/notes.tex
Added stuff on experiement with dora world

This experience is similar to the simple experiment run in \autoref{sec:simple-experiment}.
But this time using a more realistic distribution (\autoref{tab:dora-world-distribution}) with
11 different room categories and with a feature space containing semantic data for:
room size, room shape, visual appearance and 6 different objects.
Also, the features are statistically independent from each other given the room category.

The system objective was to detect novel data from a given known set of 6 categories.
Two tests where run:
\begin{enumerate}
\item Using perfect knowledge of the probability functions both thresholds where compared for performance (\autoref{fig:dora-data-perfect}).
\item The models described in \autoref{sec:dora-models} were used to learn the probability functions and the thresholds were compared (\autoref{fig:dora-data-simple}).
\end{enumerate}

\note{The graphics are a bit "cheated", since theres an added epsilon on the returned value
      to the probabilities functions to not return 0. This is because at the moment if on the known
      samples we never see a given value we don't know it belongs to the possible feature space.
      Has to be corrected.}

As possible to see in \autoref{fig:dora-data-perfect} if we are able to use the unlabelled data to build a correct $P(x)$ then we obtain an optimal detector.

\autoref{fig:dora-data-simple} shows how close our solution approaches the optimal by using creating models to for the conditional and unconditional probability from samples.

Important to notice that our model of unconditional probability is very simple, it assumes semantic data to be independent from each other.

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{dora-data/roc-perfect.pdf}
\caption{ROC curve for thresholds based on perfect information running on the dora distribution}
\label{fig:dora-data-perfect}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\textwidth]{dora-data/roc-simple.pdf}
\caption{ROC curve for thresholds learned on samples from samples in the dora distribution}
\label{fig:dora-data-simple}
\end{figure}

\begin{sidewaystable}[h]
\begin{center}
\scalebox{0.35}{\input{dora-data/explain}}
\end{center}
\caption{\label{tab:dora-world-distribution}World model used in Dora. Each column cell shows $P(feature|class)$}
\end{sidewaystable}

>>>>

This experience is similar to the simple experiment run in \autoref{sec:simple-experiment}.
But this time using a more realistic distribution (\autoref{tab:dora-world-distribution}) with
11 different room categories and with a feature space containing semantic data for:
room size, room shape, visual appearance and 6 different objects.

As expected, having perfect information of the distributions $P(x|c \in known)$ and $P(x)$
allows us to build an optimal detector as seen on \autoref{fig:dora-data-perfect}.

\begin{figure}[h]
\centering
\includegraphics[width=0.75\textwidth]{dora-data/roc-perfect.pdf}
\caption{ROC curve for thresholds based on perfect information running on the dora distribution.
         The convex curve for $P(x|k)/P(x)$ shows its optimal.}
\label{fig:dora-data-perfect}
\end{figure}

Nonetheless to build that optimal detector we need information on $P(x)$ which often is not available.
We focus then on the study of the bias introduced by using $P(x|k)$ as a threshold.
For that we compared all roc curves based on perfect information for each missing
classes (\autoref{fig:roc-curves-missing-1}).

\begin{figure}[h]
 \centering
 \subfloat[Anteroom]{\includegraphics[width=0.33\textwidth]{dora-data/roc-missing-anteroom.pdf}}
 \subfloat[Bathroom]{\includegraphics[width=0.33\textwidth]{dora-data/roc-missing-bathroom.pdf}}
 \subfloat[Computer Lab]{\includegraphics[width=0.33\textwidth]{dora-data/roc-missing-computerlab.pdf}} \\
 \subfloat[Conference Hall]{\includegraphics[width=0.33\textwidth]{dora-data/roc-missing-conferencehall.pdf}}
 \subfloat[Double Office]{\includegraphics[width=0.33\textwidth]{dora-data/roc-missing-doubleoffice.pdf}}
 \subfloat[Hallway]{\includegraphics[width=0.33\textwidth]{dora-data/roc-missing-hallway.pdf}} \\
 \subfloat[Kitchen]{\includegraphics[width=0.33\textwidth]{dora-data/roc-missing-kitchen.pdf}}
 \subfloat[Meeting Room]{\includegraphics[width=0.33\textwidth]{dora-data/roc-missing-meetingroom.pdf}}
 \subfloat[Professors Office]{\includegraphics[width=0.33\textwidth]{dora-data/roc-missing-professorsoffice.pdf}} \\
 \subfloat[Robot Lab]{\includegraphics[width=0.33\textwidth]{dora-data/roc-missing-robotlab.pdf}}
 \subfloat[Single Office]{\includegraphics[width=0.33\textwidth]{dora-data/roc-missing-singleoffice.pdf}}

\caption{\label{fig:roc-curves-missing-1}ROC curves for all one single missing category in the known data.}
\end{figure}

\note{TODO: Look into the distribution and explain why the harder new room categories to detect were
      Computer Lab, Conference Hall, Meeting Room and Robot Lab.}

\note{The graphics are a bit "cheated", since theres an added epsilon on the returned value
      so the probabilities functions do not return 0. This is because at the moment if on the known
      samples we never see a given value we don't know it belongs to the possible feature space.
      Has to be corrected.}

\subsubsection{Approximation Models}

In this section, I should describe how we use graphical models to model both $P(x|k)$ and $P(x)$.
And made them as $P(G)$ and $P(G')$.

Comparison of the estimated models with respect to the perfect information functions as been done on
\autoref{fig:dora-data-simple}.
Important to notice that our model of unconditional probability is very simple, still is able to improve
our novelty detection.

@@

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{dora-data/roc-simple.pdf}
\caption{ROC curve for thresholds learned on samples from samples in the dora distribution}
\label{fig:dora-data-simple}
\end{figure}

\begin{sidewaystable}[h]
\begin{center}
\scalebox{0.50}{\input{dora-data/explain}}
\end{center}
\caption{\label{tab:dora-world-distribution}World model used in Dora. Each column cell shows $P(feature|class)$}
\end{sidewaystable}
